---
title:      "A small R library which computes precision of descriptive statistics from measurement precision"
subtitle:   'A companion to "How many decimals?"'
author:     "Denis Cousineau"
abstract:   This user manual shows how to round commonly-used descriptive statistics from the measurement precision. It accompanies the article "How many decimals? Rounding descriptive statistics based on measurement precision", \textit{Journal of Mathematical Psychology} (submitted).
date:       "10/01/2020"
output:     pdf_document
---

<!-- to compile this document outside of RStudio, run
  setwd("C:\\Users\\DenisCousineau\\Repositories\\MeasurementPrecision-gitted")
  rmarkdown::render('UserManual_v0.22.Rmd')
-->

# Loading the library

The exact computations to round descriptive statistics are found in the main text under Table 1. The equations are all simple; however, there are a lot of formulas to remember. To assist in rounding descriptive statistics, I have designed a small R library called \texttt{MeasurementPrecision} (no space and two capital letters) which resides on GitHub. To upload it, first install the \texttt{devtools} library from CRAN (Wickham, Hester & Chang, 2019). Then prior to the first use, issue the commands

```{r setupA, eval=FALSE}
devtools::install_github("dcousin3/MeasurementPrecision")
library(MeasurementPrecision)
```

On subsequent sessions, you can only use

```{r setupB, include = TRUE}
library(MeasurementPrecision)
```

# A basic use

Let's assume the following two sets of observations

```{r init, include = TRUE}
sample1 <- c(83, 58, 79, 50, 49, 53, 62, 79, 66)
sample2 <- c(71, 62, 83, 93, 56, 82, 66, 69, 82, 86, 74, 61, 59, 101, 94, 86, 75)
```

To get a rounded descriptive statistic, use a command named \texttt{roundMP.\textit{statistic}}. For example, to round the mean of the first sample, use:

```{r test1a, include = TRUE, eval = FALSE}
roundMP.mean(fromData = sample1, deltax = 0.5)
```
    
where \texttt{deltax}, a mandatory argument, is the precision of the instrument. The command returns a one-line data frame with four columns:

```{r test1b, include = TRUE, echo = FALSE, comment = ""}
roundMP.mean(fromData = sample1, deltax = 0.5)
```

where \texttt{machine.precision} is the unrounded result, \texttt{extrinsic} is the  extrinsic precision-based rounding; \texttt{systematic} is the result assuming systematic measurement error and \texttt{non.systematic} is the result assuming non-systematic measurement error.

In any of the commands, you can use \texttt{fromData} if you want to specify raw data or \texttt{fromStatistics} to provide already-calculated descriptive statistics (provide them with all the precision you can). For example, 

```{r test2, include = TRUE, eval = FALSE, comment = ""}
roundMP.mean(
	fromStatistics = list(mean = 64.333333, sd = 13.20982, n = 9), 
	deltax = 0.5
) 
```

returns the same results as above. If you issue this command with an empty list of statistics, an error message will let you know which statistics are required.

# Getting rounded statistics beyond the mean

You can also round the standard deviation (\texttt{sd}), the standard error of the mean (\texttt{semean}) and the confidence interval of the mean (\texttt{cimean}): 

```{r test3, include = TRUE, echo = TRUE, collapse = FALSE, comment = ""}
roundMP.sd(fromData = sample1, deltax = 0.5)
roundMP.semean(fromData = sample1, deltax = 0.5)
roundMP.cimean(fromData = sample1, deltax = 0.5) 
```

In \texttt{roundMP.cimean}, add \texttt{gamma = } for a different coverage. For example, \texttt{gamma = 0.80} will round a 80% confidence interval of the mean.

A one-sample t-test requires the null hypothesis for the mean, provided with \texttt{mu0}, for example:

```{r test4, include = TRUE, echo = TRUE, collapse = FALSE, comment = ""}
roundMP.t.test(fromData = sample2, mu0 = 65, deltax = 0.5)
````

where 65 kg is the average planetary body weight for humans.

For statistics on two independent samples, you can use

```{r test5, include = TRUE, echo = TRUE, collapse = FALSE, comment = ""}
roundMP.meandiff(fromData = list(sample2, sample1), deltax = 0.5)
roundMP.sdpool(fromData = list(sample2, sample1), deltax = 0.5)
roundMP.cohen.d(fromData = list(sample2, sample1), deltax = 0.5)
```

The argument \texttt{fromData} accepts vectors, matrices, data frames or a list of vectors (as illustrated here).

The non-systematic estimates are by default obtained from the simplifying assumptions described in Appendix B of the main paper. To use the full expression (non-parametric solution), add \texttt{assumptions = FALSE} to any of the commands, for example

```{r test6, include = TRUE, echo = TRUE, collapse = FALSE, comment = ""}
roundMP.t.test(
    fromData = list(sample2, sample1), 
	deltax=0.5, assumptions = FALSE
)
```

Generally, there is not much differences whether the simplified or the full expression are used.

# Specifying the desired scenario as default

By default, all the roundMP functions output 4 different results, following four approaches to rounding: "\texttt{machine.precision}", "\texttt{extrinsic}", "\texttt{systematic}", and "\texttt{non.systematic}". It is possible to select only one or a few scenarios among this list by setting the global option \texttt{roundMP.selectedScenario}. For example, the following will display results only for two scenario (rounding assuming extrinsic and on systematic measurement error):

```{r test7, include = TRUE, echo = TRUE, collapse = FALSE, comment = ""}
options(roundMP.selectedScenario = c("extrinsic", "systematic"))
roundMP.mean(fromData = sample1, 0.5)
```

```{r test7off, include = TRUE, echo = FALSE}
options(roundMP.selectedScenario = c("machine.precision", "extrinsic", "systematic", "non.systematic"))
```

# Arguments \texttt{fromData}, \texttt{fromStatistics} and \texttt{fromObject}

Regarding \texttt{t.test}, it is also possible to get a rounded result from a \texttt{t.test} object directly using the argument \texttt{fromObject} instead (instead of \texttt{fromData} or \texttt{fromStatistics}). The input has to contain a t-test, not a Welch test (so use var.equal = TRUE for two-samples): 

```{r test8, include = TRUE, echo = TRUE, collapse = FALSE, comment = ""}
res <- t.test(sample1, sample2, var.equal = TRUE)
roundMP.t.test(fromObject = res, deltax = 0.5)
```

Note that the library \texttt{MeasurementPrecision} must be loaded prior to use the \texttt{t.test} function as the t.test function is redefined by \texttt{MeasurementPrecision}.

# Detailed output

Finally, to obtain more details on the computations, and see the exact precision, you can add the option \texttt{verbose = TRUE} to any command. For example, 

```{r test9, include = TRUE, echo = TRUE, collapse = FALSE, comment = ""}
roundMP.mean(fromData = list(sample1), deltax = 0.5, verbose = TRUE)
```

The results of the last line are identical to the solution provided earlier, but detailed information preceeds the results, showing the exact precision for each scenario, whether a simplifying assumption was used, and the resulting rounded result.




